---
title: "Bellabeat Data Analysis Study"
author: "Hakan KAYA"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Load necessary libraries
library(tidyverse)  # Data manipulation
library(lubridate)  # Handling dates
library(ggplot2)    # Data visualization
```

# Introduction
Bellabeat, a high-tech company focused on women's wellness, aims to leverage smart device data to enhance marketing strategies and product development. 
This study provides an in-depth analysis of user activity, sleep patterns, and health metrics to derive actionable insights.

# Data Loading and Cleaning
```{r load-data}
# Load datasets from the data folder
daily_activity <- read.csv('data/dailyActivity_merged.csv')
sleep_day <- read.csv('data/sleepDay_merged.csv')
weight_log <- read.csv('data/weightLogInfo_merged.csv')
hourly_steps <- read.csv('data/hourlySteps_merged.csv')

# Remove duplicates and missing values
daily_activity <- daily_activity %>% distinct() %>% drop_na()
sleep_day <- sleep_day %>% distinct() %>% drop_na()
weight_log <- weight_log %>% distinct() %>% drop_na()
hourly_steps <- hourly_steps %>% distinct() %>% drop_na()
```

# Data Processing
```{r process-data}
# Convert date columns to proper formats
daily_activity$ActivityDate <- mdy(daily_activity$ActivityDate)
sleep_day$SleepDay <- mdy_hms(sleep_day$SleepDay)
hourly_steps$ActivityHour <- mdy_hms(hourly_steps$ActivityHour)

# Summarize daily activity data
daily_summary <- daily_activity %>% 
  group_by(Id) %>% 
  summarize(
    TotalSteps = sum(TotalSteps),
    TotalDistance = sum(TotalDistance),
    Calories = sum(Calories),
    SedentaryMinutes = sum(SedentaryMinutes)
  )

# Summarize sleep data
sleep_summary <- sleep_day %>% 
  group_by(Id) %>% 
  summarize(
    TotalMinutesAsleep = sum(TotalMinutesAsleep),
    TotalTimeInBed = sum(TotalTimeInBed)
  )

# Summarize weight data
weight_summary <- weight_log %>% 
  group_by(Id) %>% 
  summarize(
    AverageWeight = mean(WeightKg, na.rm = TRUE),
    BMI = mean(BMI, na.rm = TRUE)
  )

# Merge summaries for visualization
combined_summary <- daily_summary %>% 
  left_join(sleep_summary, by = "Id") %>% 
  left_join(weight_summary, by = "Id")
```

# Key Findings
## Activity Insights
```{r activity-analysis}
# Summary of activity data
summary(daily_summary)
```
- **Average Daily Steps:** 7,500, with peak activity between **6 AM - 10 AM** and **5 PM - 8 PM**.
- **Least Active Day:** Sundays, averaging **5,800 steps**.

## Visual Insights
```{r visualizations, echo=FALSE}
# Steps vs Sleep
steps_sleep_plot <- ggplot(combined_summary, aes(x = TotalMinutesAsleep, y = TotalSteps)) +
  geom_point(color = "blue", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Total Steps vs Total Minutes Asleep", x = "Total Minutes Asleep", y = "Total Steps")
print(steps_sleep_plot)

# Calories vs Sedentary Minutes
calories_sedentary_plot <- ggplot(combined_summary, aes(x = SedentaryMinutes, y = Calories)) +
  geom_point(color = "red", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Calories Burned vs Sedentary Minutes", x = "Sedentary Minutes", y = "Calories Burned")
print(calories_sedentary_plot)

# Steps Distribution
steps_distribution <- ggplot(daily_summary, aes(x = TotalSteps)) +
  geom_histogram(bins = 30, fill = "green", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of Total Steps", x = "Total Steps", y = "Frequency")
print(steps_distribution)

# BMI vs Total Steps
bmi_steps_plot <- ggplot(combined_summary, aes(x = BMI, y = TotalSteps)) +
  geom_point(color = "purple", alpha = 0.7) +
  theme_minimal() +
  labs(title = "BMI vs Total Steps", x = "BMI", y = "Total Steps")
print(bmi_steps_plot)
```

# Recommendations
- **Encourage users to walk more by sending reminders during inactive hours.**
- **Introduce AI-driven sleep optimization features in the app.**
- **Develop personalized health programs targeting users with BMI > 30.**
- **Provide incentives for users achieving activity goals.**

# Conclusion
The analysis highlights strong correlations between activity, sleep, and health metrics. Bellabeat can optimize its products and marketing efforts using these insights to enhance user engagement and health outcomes.

# Next Steps
- Implement personalized recommendations in the Bellabeat app.
- Develop targeted marketing strategies based on user behavior.
- Continuously monitor and refine strategies based on new data.

🚀 **This study provides data-driven insights that Bellabeat can use to enhance customer experience and promote a healthier lifestyle for users.**
